## JHU: R Programming
### Programming Assignment 2

This programming assignment investigates lexical scoping and environments in R, particularly when 
used in the context of caching output from costly operations on data.

The assignment begins with a special vector class definition:

```{r}
makeVector <- function(x = numeric()) {
    m <- NULL
    set <- function(y) {
        x <<- y
        m <<- NULL
    }
    get <- function() x
    setmean <- function(mean) m <<- mean
    getmean <- function() m
    list(set = set, get = get,
         setmean = setmean,
         getmean = getmean)
}
```

Using a list, a vector of numeric atoms passed to the `makeVector` function creates a list to wrap the original vector as well as a container for storing the mean of the vector's values. 

Note the use of `<<-` for assignment. With lexical scoping symbols are resolved from the environment where the fuction was defined. The parent environment is where the function was *called* rather than where it was *defined*. The `<<-` specifically asks the function to look in the parent environment rather than the environment where the function was defined.

As an example of this vector:

```{r}
aVector <- makeVector(c(1, 2, 3, 4, 5))
```

The line `get <- function() x` means that calling <vector>$get() will simply return the vector unaltered. The vector is accessed like a part of a list

```{r}
aVector$get()
```

### Caching the Mean

An example of how to cache an operation in provided with

```{r}
cachemean <- function(x, ...) {
        m <- x$getmean()
        if(!is.null(m)) {
                message("getting cached data")
                return(m)
        }
        data <- x$get()
        m <- mean(data, ...)
        x$setmean(m)
        m
}
```

One of the elements of a makeVector list is `m` though it is initialized as NULL. This new function `cachemean`, returns the value of `m` *if it is set*. If `m` is not set (it is NULL), then `cachemean` retrieves the vector from the list, calculates the mean, puts the mean into `m`, and returns the mean. The next time `cachemean` is called on the makeVector list/object a value in `m` will be available. These subsequent calls will be much faster than the first one as shown by the calculation of the mean of a vector with 100 million elements.

```{r}
bVector <- makeVector(sample(x=20, size=100e6, replace = TRUE))
system.time(cachemean(bVector))
system.time(cachemean(bVector))
```

### Cachable Matrix
####Creating matrices with containers for caching results of costly computations

First we create the matrix container, a list, like the array container above.

```{r}
makeCacheMatrix <- function(x = numeric()) {
    s <- NULL
    set <- function(y) {
        x <<- y
        s <<- NULL
    }
    get <- function() x
    setinv <- function(solve) s <<- solve
    getinv <- function() s
    list(set = set, get = get,
         setinv = setinv,
         getinv = getinv)
}
```

Instantiating a makeCacheMatrix stores the passed matrix in the environment so that 
it is returned with `<matrix>$get()`. This definition also includes a container `s` 
for storing the inverse of the matrix.

Next there is a function for retrieving the inverse, or if the inverse has not yet 
been calculated and cannot be returned, it is calculated, stored, and then returned.

```{r}
cacheInverse <- function(x, ...) {
        s <- x$getinv()
        if(!is.null(s)) {
                message("getting cached data")
                return(s)
        }
        data <- x$get()
        s <- solve(data, ...)
        x$setinv(s)
        s
}
```

To try it out, we can create a cachable matrix that has 4 million elements.

```{r}
set.seed(20)
aMatr <- makeCacheMatrix(matrix(
    sample(x = 100, size = 2000^2, replace = TRUE), ncol = 2000))
```

And try then find its inverse:

```{r}
system.time(cacheInverse(aMatr))
system.time(cacheInverse(aMatr))
```

